New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2022-09-24 paper/appendix.Rmd
I fixed some typos

The problem with new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.

@@

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$, formed by its first two columns. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${x_{m1},x_{m2}}\to {a^*_{m1},a^*_{m2}}$
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuouns and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.

\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\end{align}

@@

{\textcolor{blue} Do we need three methods? is this one implemented somewhere?}

@@

The function renders a manual tour with controls that allow the user to navigate through different projections. For a given projection, a slider permits the user to change the slide width within its range and a separate box switches the view to a projection. The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders. An additional box displays the coordinates of the projection matrix explicitly.

@@

>>>>

The problem with the new simple method (Algorithm 1) is that the precise values for $V_m$ cannot be specified because the orthonormalisation will change them.

@@

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuous  and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.

\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\label{matrices}
\end{align}

@@

{\textcolor{blue} Ursula, I am more convinced now that this is exactly method 2, the only difference is that you want to rotate to get to the "first" direction, whereas I simply relabel the rows of the matrix}

@@

The function renders a manual tour with controls that allow the user to navigate through different projections. For a given projection, a slider permits the user to change the slice width within its range and a separate box switches the view to a projection. The user can also change the centre point of slices; the size of the points rendered and the scale of the plot through sliders. An additional box displays the coordinates of the projection matrix explicitly.

@@

\textcolor{red}{does this mean that we are changing the name of the package from mmtour.wl to Slice Tour notebook?}

### 2022-09-24 paper/appendix.tex
I fixed some typos

The problem with new simple method (Algorithm 1) is that the precise
values for \(V_m\) cannot be specified because the orthonormalisation
will change them.

@@

For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
of \(O\), formed by its first two columns. Whereas orthonormality of the
basis for the p-dimensional space is given by
\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
projection matrix is expressed as
\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
to guarantee orthonormality of the new projection matrix. One possible
algorithm to achieve this is

\begin{enumerate}
\item Cursor movement takes ${x_{m1},x_{m2}}\to {a^*_{m1},a^*_{m2}}$
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}

The random completion of \(O\) outside the projection matrix provides an
exploration of dimensions orthogonal to the projection plane. However,
it renders projections in a way that is not continuouns and may be
distracting. This can be alleviated by replacing the random completion
with a rule restricting the size of the jumps, for example in step (b)
the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
orthogonalisation.

\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\end{align}

@@

\{\textcolor{blue} Do we need three methods? is this one implemented
somewhere?\}

>>>>

The problem with the new simple method (Algorithm 1) is that the precise
values for \(V_m\) cannot be specified because the orthonormalisation
will change them.

@@

For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
of \(O\) (an orthonormal basis for the p-dimensional space), formed by
its first two columns as illustrated in
Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
for the p-dimensional space is given by
\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
projection matrix is expressed as
\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
to guarantee orthonormality of the new projection matrix. One possible
algorithm to achieve this is

\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}

The random completion of \(O\) outside the projection matrix provides an
exploration of dimensions orthogonal to the projection plane. However,
it renders projections in a way that is not continuous and may be
distracting. This can be alleviated by replacing the random completion
with a rule restricting the size of the jumps, for example in step (b)
the values \(a_{i\neq m,j\geq3}\) can be left unchanged before
orthogonalisation.

\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\label{matrices}
\end{align}

@@

\{\textcolor{blue} Ursula, I am more convinced now that this is exactly
method 2, the only difference is that you want to rotate to get to the
``first'' direction, whereas I simply relabel the rows of the matrix\}

### 2022-09-24 paper/paper.Rmd
I fixed some typos

In addition, we also want to be able to look at slices of the data and select slicing parameters interactively. Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice height $h$ and a numeric input to specify the center point $c$ explicitly. The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).

@@

The most relevant inbuilt Mathematica function for our purpose is the locator pane. This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.

The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group). After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visulisation appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre. The user can also change the appearance of the plot by zooming into the center or changing the point size with sliders provided. It is worth noting that this zoom works best with data that has been centred and scaled. The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration later, as shown in the applications below.

@@

Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice similar to how the manual tour was implemented here. An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.

@@

We start by first exploring the projections of the model prediction. Figure \ref{proj1} summarises the process. Through manual rotation of the view we can get a feeling for where in the space we primarily predict each of the three species, and we also get a sense of the difference between the two models. To illustrate this difference we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model. This particular projection ($A_1$) is exported so that the same projection can be used to show the LDA model (middle plot) and the actual data (right plot). What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes. And, interestingly this particular projection of the original data shows very distinct clusters of the three species. That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.

@@

We continue the investigation by now slicing orthogonally to  the projection $A_1$. For both models we look at a thin ($h=0.5$) slice through the center, $\mathbf{c}^0 = (0,0,0,0)$. At  first we explore how changing the projection away from $A_1$ can help with understanding the boundary better. For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.

@@

Finally, to determine which model does the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}). Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group. It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data. The LDA boundary is better placed for all species.

>>>>

In addition, we also want to be able to look at slices of the data and select slicing parameters interactively. Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness $h$ (dubbed `height` in the `Mathematica` package) and a numeric input to specify the center point $c$ explicitly. The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).

@@

The most relevant inbuilt Mathematica function for our purpose is the `LocatorPane`. This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour.

The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group). After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualisation appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice centre. The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided. It is worth noting that this zoom works best with data that has been centred and scaled. The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration into a later study, as shown in the applications below.

@@

Potential adjustments for future work include a better visualization of the current center point, potentially with interactive graphics to move the slice, similarly to how the manual tour was implemented here. An alternative would be a step-wise shift of the slice towards the new center point to get a better understanding of the distribution along the selected direction.

@@

We start by exploring the projections of the model prediction. Figure \ref{proj1} summarises the process. Rotating manually the view, we can visualise the location of predictions for each of the three species, and also get a sense of the difference between the two models. 
To illustrate this difference, we have manually rotated the projection for the RF model (left plot) to identify a view that shows the non-linear but block-type structure that is typical of this type of model. This particular projection ($A_1$) is exported so that it can later be used to show the LDA model (middle plot) and the actual data (right plot). What can be seen is the linearity of the LDA model, where the boundaries are linear and oblique to the variable axes. And, interestingly, this particular projection of the original data shows very distinct clusters of the three species. That means, the obscuring of the boundaries between groups for both of the models is driven by what is happening in the orthogonal space to the plane of the selected projection.

@@

We now continue the investigation by slicing orthogonally to  the projection $A_1$. For both models we look at a thin ($h=0.5$) slice through the center, $\mathbf{c}^0 = (0,0,0,0)$. At  first we explore how changing the projection away from $A_1$ can help with understanding the boundary better. For our example, notice that $A_1$ does not contain any contribution from the second variable (`bd`), so we will first rotate this variable into the view.

@@

Finally, to determine which model describes the boundaries better, compare them with the $A_2$ projection of the data (Figure \ref{proj2}). Gentoo is more separated from the other two in this projection, and one can imagine that the trees in the forest has greedily grasped any one of many places to make a split to separate the group. It might be argued though the that RF boundary is cut too close to the Chinstrap species, and might lead to some unnecessary misclassification with new data. The LDA boundary is better placed for all species.

### 2022-09-24 paper/paper.tex
I fixed some typos

In addition, we also want to be able to look at slices of the data and
select slicing parameters interactively. Here we have implemented a
switch to change between projection and slice view, a slider that
adjusts the slice height \(h\) and a numeric input to specify the center
point \(c\) explicitly. The display will update based on these inputs,
in particular the view will jump to a new slice when \(c\) is changed
(though an interpolation might also be useful).

@@

The most relevant inbuilt Mathematica function for our purpose is the
locator pane. This creates a region on the screen where the position of
the mouse is captured and then converted to input that updates the
graphics functions, enabling the manual navigation of the tour.

The examples presented in the application section below (and attached in
notebook format in the supplementary material) all use our primary new
function, \texttt{SliceDynamic}. This function typically accepts grouped
data in the form of a matrix where the second last column details the
name of the group and the last column details the group index (which can
be one if there is only one group). After specifying the initial slice
thickness and the slice range, the user is presented with an interactive
display in which the control objects appear on the left and the slice
visulisation appears on the right. The user can change the orientation
of the slice via the locator pane, which changes the projection matrix;
a slider controls the slice thickness; and there is an input field,
which changes the slice centre. The user can also change the appearance
of the plot by zooming into the center or changing the point size with
sliders provided. It is worth noting that this zoom works best with data
that has been centred and scaled. The projected data can be displayed to
contrast it with the slice by ticking a box. The explicit projection
matrix can also be displayed via another checkbox. This is especially
useful when we wish to import a projection that was identified as
interesting in the manual exploration later, as shown in the
applications below.

@@

We start by first exploring the projections of the model prediction.
Figure \ref{proj1} summarises the process. Through manual rotation of
the view we can get a feeling for where in the space we primarily
predict each of the three species, and we also get a sense of the
difference between the two models. To illustrate this difference we have
manually rotated the projection for the RF model (left plot) to identify
a view that shows the non-linear but block-type structure that is
typical of this type of model. This particular projection (\(A_1\)) is
exported so that the same projection can be used to show the LDA model
(middle plot) and the actual data (right plot). What can be seen is the
linearity of the LDA model, where the boundaries are linear and oblique
to the variable axes. And, interestingly this particular projection of
the original data shows very distinct clusters of the three species.
That means, the obscuring of the boundaries between groups for both of
the models is driven by what is happening in the orthogonal space to the
plane of the selected projection.

@@

We continue the investigation by now slicing orthogonally to the
projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first we explore
how changing the projection away from \(A_1\) can help with
understanding the boundary better. For our example, notice that \(A_1\)
does not contain any contribution from the second variable
(\texttt{bd}), so we will first rotate this variable into the view.

@@

Finally, to determine which model does the boundaries better, compare
them with the \(A_2\) projection of the data (Figure \ref{proj2}).
Gentoo is more separated from the other two in this projection, and one
can imagine that the trees in the forest has greedily grasped any one of
many places to make a split to separate the group. It might be argued
though the that RF boundary is cut too close to the Chinstrap species,
and might lead to some unnecessary misclassification with new data. The
LDA boundary is better placed for all species.

>>>>

In addition, we also want to be able to look at slices of the data and
select slicing parameters interactively. Here we have implemented a
switch to change between projection and slice view, a slider that
adjusts the slice thickness \(h\) (dubbed \texttt{height} in the
\texttt{Mathematica} package) and a numeric input to specify the center
point \(c\) explicitly. The display will update based on these inputs,
in particular the view will jump to a new slice when \(c\) is changed
(though an interpolation might also be useful).

@@

The most relevant inbuilt Mathematica function for our purpose is the
\texttt{LocatorPane}. This creates a region on the screen where the
position of the mouse is captured and then converted to input that
updates the graphics functions, enabling the manual navigation of the
tour.

The examples presented in the application section below (and attached in
notebook format in the supplementary material) all use our primary new
function, \texttt{SliceDynamic}. This function typically accepts grouped
data in the form of a matrix where the second last column details the
name of the group and the last column details the group index (which can
be one if there is only one group). After specifying the initial slice
thickness and the slice range, the user is presented with an interactive
display in which the control objects appear on the left and the slice
visualisation appears on the right. The user can change the orientation
of the slice via the locator pane, which changes the projection matrix;
a slider controls the slice thickness; and there is an input field,
which changes the slice centre. The user can also change the appearance
of the plot by zooming into the center or changing the point size with
the sliders provided. It is worth noting that this zoom works best with
data that has been centred and scaled. The projected data can be
displayed to contrast it with the slice by ticking a box. The explicit
projection matrix can also be displayed via another checkbox. This is
especially useful when we wish to import a projection that was
identified as interesting in the manual exploration into a later study,
as shown in the applications below.

@@

We start by exploring the projections of the model prediction. Figure
\ref{proj1} summarises the process. Rotating manually the view, we can
visualise the location of predictions for each of the three species, and
also get a sense of the difference between the two models. To illustrate
this difference, we have manually rotated the projection for the RF
model (left plot) to identify a view that shows the non-linear but
block-type structure that is typical of this type of model. This
particular projection (\(A_1\)) is exported so that it can later be used
to show the LDA model (middle plot) and the actual data (right plot).
What can be seen is the linearity of the LDA model, where the boundaries
are linear and oblique to the variable axes. And, interestingly, this
particular projection of the original data shows very distinct clusters
of the three species. That means, the obscuring of the boundaries
between groups for both of the models is driven by what is happening in
the orthogonal space to the plane of the selected projection.

@@

We now continue the investigation by slicing orthogonally to the
projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first we explore
how changing the projection away from \(A_1\) can help with
understanding the boundary better. For our example, notice that \(A_1\)
does not contain any contribution from the second variable
(\texttt{bd}), so we will first rotate this variable into the view.

@@

Finally, to determine which model describes the boundaries better,
compare them with the \(A_2\) projection of the data (Figure
\ref{proj2}). Gentoo is more separated from the other two in this
projection, and one can imagine that the trees in the forest has
greedily grasped any one of many places to make a split to separate the
group. It might be argued though the that RF boundary is cut too close
to the Chinstrap species, and might lead to some unnecessary
misclassification with new data. The LDA boundary is better placed for
all species.
